# Question 1:

-   I have chosen the plot of avg.height and weight vs summer sports. For this plot in my project we received feedback requiring a bit more clarification from this plot.

-   Hence for the original plot to be more clear, I have reduced the number of sports but still maintained the trend that was being followed.

-   The reducing mean height and weight can be used to classify which player will be better suited for which sport.

```{r}
olympics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-08-06/olympics.csv')
```

```{r}
library(pacman)

# Load necessary packages
pacman::p_load(ggplot2, dplyr, ggrepel, gganimate, png, magick, ggridges)
```

```{r}
#| label: avg-height-by-sport 

# Filter for Summer Olympics
summer <- dplyr::filter(olympics, season == "Summer")  |> na.omit(summer$height)

# Calculate average height while removing NA values
summer_height_avg <- tapply(summer$height, summer$sport, function(x) mean(x, na.rm = TRUE))

# Convert to data frame
summer_height_avg_df <- data.frame(sport = names(summer_height_avg), average_height = summer_height_avg)

# Sort by average height
summer_height_avg_df <- summer_height_avg_df[order(-summer_height_avg_df$average_height), ] 

# Print the result
summer$sport <- factor(summer$sport, levels = summer_height_avg_df$sport[order(summer_height_avg_df$average_height)])
summer_clean <- summer |> na.omit()
sport_counts <- table(summer_clean$sport)
sports_with_enough_data <- names(sport_counts[sport_counts >= 750])
# Create the violin plot
summer |> filter(sport %in% sports_with_enough_data) |>
ggplot(
  aes(x = height, y = sport, fill = sex)) +
   geom_density_ridges(alpha = 0.3) +
  labs(title = "Average Height Distribution by Sport and Season in the Olympics",
       x = "Height",
       y = "Sport") +
  theme_minimal()
```

```{r}
#| label: avg-weight-by-sport

summer <- dplyr::filter(olympics, season == "Summer") |> na.omit(summer$weight)
sport_counts <- summer %>%
  group_by(sport) %>%
  summarise(count = n())

valid_sports <- sport_counts %>%
  filter(count >= 750) %>%
  pull(sport)

summer_filtered <- summer %>%
  filter(sport %in% valid_sports)

# Calculate average weight while removing NA values
summer_weight_avg <- tapply(summer_filtered$weight, summer_filtered$sport, function(x) mean(x, na.rm = TRUE))

# Convert to data frame
summer_weight_avg_df <- data.frame(sport = names(summer_weight_avg), average_weight = summer_weight_avg)

# Sort by average weight
summer_weight_avg_df <- summer_weight_avg_df[order(-summer_weight_avg_df$average_weight), ] 

# Print the result
summer_filtered$sport <- factor(summer_filtered$sport, levels = summer_weight_avg_df$sport[order(summer_weight_avg_df$average_weight)])
# Create the violin plot
summer_filtered |>
ggplot(
  aes(x = weight, y = sport, fill = sex)) +
  geom_density_ridges(alpha = 0.3) +
  labs(
    title = "Average Weight Distribution by Sport and Season in the Olympics",
    x = "Weight",
    y = "Sport") +
  theme_minimal()
```

# Question 2:

```{r}
library(tigris)
library(sf)
library(ggrepel)
```

```{r}
options(tigris_use_cache = TRUE)

az_counties <- tigris::counties("Arizona", year = 2021, progress_bar = FALSE)

write_sf(az_counties, "Arizona_Counties_2021.shp")
```

```{r}
az_counties_check <- read_sf("Arizona_Counties_2021.shp")
head(az_counties_check)
```

```{r}
ggplot(data = az_counties_check)+
  geom_sf(fill = "#ececec", color = "#949494")+
  geom_label_repel(data = az_counties_check,
                    aes(label = NAME, geometry = geometry), stat = "sf_coordinates",
                   min.segment.length = 0)+
  labs(
    title = "Counties in Arizona State",
    x = "Longitude",
    y = "Latitude",
    caption = "Source: Shapefile obtained using {tigris}"
  )+
  theme_minimal()

```

# Question 3:

```{r}
library(readxl)
library(stringr)
```

```{r}
# Load the shapefile data for Arizona
arizona_counties <- counties(state = "AZ", cb = TRUE, year = 2023)

# Load the data and inspect the column structure
population_data <- read_excel(
  "C:/Users/ual-laptop/Downloads/Homework 4  attached files Nov 27, 2024 728 PM/HW4_data/HW4_data/co-est2023-pop-04.xlsx",
  skip = 2      
)

# Rename columns for easier handling
colnames(population_data) <- c(
  "geographic_area", "april_2020_estimates", "pop_2020", 
  "pop_2021", "pop_2022", "pop_2023"
)

# Filter out state-level data and keep only county-level rows
population_data <- population_data %>%
  filter(!grepl("^Arizona", geographic_area)) %>%
  mutate(
    county = gsub(" County, Arizona", "", geographic_area),  # Clean up county names
    total_pop_change_20_23 = pop_2023 - pop_2020            # Calculate population change
  )

# Select relevant columns
population_data <- population_data %>%
  select(county, total_pop_change_20_23)
population_data <- population_data %>% slice(-1) %>%    # Remove first row
  slice_head(n=15)
population_data <- population_data %>%
  mutate(county = str_remove(county,"^\\."))
```

```{r}
# Merge the shapefile data with population change data
arizona_counties <- arizona_counties %>%
  left_join(population_data, by = c("NAME" = "county"))
# Plot the map with population change
ggplot(data = arizona_counties) +
  geom_sf(aes(fill = total_pop_change_20_23), color = "White", lwd = 0.2) +
  scale_fill_distiller(
    palette = "RdBu",
    direction = -1,
    name = "Population change"
  ) +
  labs(
    title = "Resident Population Change for Counties in AZ",
    subtitle = "July 01, 2020 to July 01, 2023",
    x = "Longitude",
    y = "Latitude",
    caption = "Source: Shapefile obtained using {tigris} R package, v2.0.1\npopulation change data from the US Census Bureau"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0, size = 14),
    plot.subtitle = element_text(hjust = 0, size = 12),
    plot.caption = element_text(hjust = 0, size = 8),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    legend.position="right",
    panel.grid.major = element_line(size=0.7)
)
```

# Question 4:

```{r}
tribal_regions <- read_sf("C:/Users/ual-laptop/Downloads/Homework 4  attached files Nov 27, 2024 728 PM/HW4_data/HW4_data/American_Indian_Reservations_in_Arizona.shp")

tribal_regions <- st_transform(tribal_regions, crs = 4269)

ggplot()+
  geom_sf(data = az_counties_check, fill = "#ececec", color = "white")+
  geom_sf(data = tribal_regions, color = "black", fill = "#d3d3d3", linewidth = 1)+
  geom_label_repel(data = head(tribal_regions), aes(label = NAME, geometry = geometry), stat = "sf_coordinates",
                    min.segment.length = 0)+
  labs(
    title = "Indigenous Tribal Boundaries in AZ",
    y = "Latitude",
    x = "Longitude",
    caption = "Source: Shapefile obtained using {tigris} R package, \n
                Indigenous Tribal shapefile obtained from AZGeoData"
  )+
  theme_minimal()
```

# Question 5:

```{r}
filtered_az_counties <- arizona_counties %>%
  filter(NAME %in% c("Maricopa", "Pinal", "Pima"))

```

```{r}
base_map <- ggplot(data = arizona_counties) +
  geom_sf(aes(fill = total_pop_change_20_23), color = "white") +  
 scale_fill_distiller(
    palette = "RdBu",
    direction = -1,
    name = "Population change"
  ) +
  theme_minimal() +
  geom_label_repel(data = filtered_az_counties,  
                   aes(label = NAME, geometry = geometry),
                   stat = "sf_coordinates", 
                   size = 3, fontface = "bold",
                   fill = "white", color = "black", 
                   label.padding = unit(0.2, "lines"),  
                   label.size = 0.3,                 
                   arrow = arrow(length = unit(0.02, "npc")),  
                   force = 5,                        
                   max.overlaps = Inf     
  ) +
  labs(
    title = "Resident Population Change for Counties in AZ",
    subtitle = "July 01, 2020 to July 01, 2023",
    x = "Longitude",
    y = "Latitude",
    caption = "Source: Shapefile obtained using {tigris} R package, v2.0.1 \npopulation change data from the US Census Bureau \nIndigenous Tribe Shapefile btained from AZGeo Data") +
  theme(
    axis.text = element_text(size = 10), 
    axis.ticks = element_line(),
    plot.title = element_text(size = 14, hjust = 0.6),
    plot.subtitle = element_text(size = 10, hjust = -0.2),
    plot.caption = element_text(size = 9, hjust = 1),
    axis.text.x = element_text(size = 8),
    legend.position = c(0.27, 0.9),       
    legend.direction = "horizontal",     
    legend.background = element_rect(fill = NA, color = NA),  
    legend.title = element_text(size = 10),  
    legend.text = element_text(size = 8),  
    legend.title.position = "top",
    legend.spacing.y = unit(0.3, "cm"),  
    legend.key.size = unit(0.7, "cm"),   
    legend.key.width = unit(0.95, "cm"),     
    legend.key.height = unit(0.4, "cm")   
        ) +
  coord_sf(xlim = c(-122, -109), ylim = c(29, 37))  
base_map
```

```{r}
filtered_tribes_sub <- tribal_regions %>%
  filter(NAME %in% c("White Mountain Apache Tribe",
                     "San Carlos Apache Tribe",
                     "Tohono O’odham Nation"))
inset_map <- ggplot(data = arizona_counties) +
  geom_sf(aes(fill = total_pop_change_20_23), color = "white") +  
  scale_fill_distiller(
    palette = "RdBu",
    direction = -1,
    name = "Population change"
  ) +  
  geom_sf(data = tribal_regions, 
          fill = "gray", color = "black", size = 1) +
  
  geom_label_repel(data = filtered_tribes_sub,   
                   aes(label = NAME, geometry = geometry),
                   stat = "sf_coordinates",  
                   size = 1.8, fontface = "bold",
                   fill = "white", color = "black", 
                   label.padding = unit(0.2, "lines"),  
                   label.size = 0.3,           
                   arrow = arrow(length = unit(0.02, "npc")),
                   force = 10,                    
                   max.overlaps = Inf               
  )+
  theme_minimal() +
  theme(
    axis.title = element_blank(),  
    axis.text = element_blank(),  
    axis.ticks = element_blank(),  
    plot.background = element_rect(fill = "grey50", color = NA), 
    panel.background = element_rect(fill = "grey50", color = NA)
  )+
  guides(fill = guide_none())
inset_map
```

```{r}
library(patchwork)

base_map_with_lines <- base_map +
  geom_segment(aes(x = -113, y = 32, xend = -118, yend = 32), linetype = "dotted", size = 0.7, color = "black") +
  geom_segment(aes(x = -113, y = 32, xend = -118, yend = 30), linetype = "dotted", size = 0.7, color = "black") +
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))  
compressed_inset <- inset_map +
  theme(plot.margin = unit(c(0.3,0.3,0.1,0.1), "cm"))  
combined_plot <- base_map_with_lines +
  inset_element(
    compressed_inset,
    left = -0.2,
    bottom = 0,
    right = .6,
    top = .5
  )
combined_plot
```
