QUESTION 1
```{r}
library(magrittr)
library(dplyr)
```


```{r}
load(file = "C:/Users/ual-laptop/Downloads/INFO526 HW2/data/edibnb.rda")
neighborhood_ordr <- edibnb %>%
  group_by(neighbourhood) %>%
  summarise(median_score = median(review_scores_rating, na.rm= TRUE)) %>%
  arrange(median_score) %>%
  pull(neighbourhood)
```


```{r}
library(ggplot2)
library(ggridges)

ggplot(edibnb, aes(x= review_scores_rating, y= factor(neighbourhood, levels = neighborhood_ordr )))+
  geom_density_ridges(fill = "lightblue", color = "white")+
  labs(title = "Ridgeplot of Airbnb review scores of Edinburgh neighborhoods",
         x= "Review Scores Rating", 
       y= "Neighbourhoods Ordered by Mean Value")
```

- This density ridge plot displays the distribution of airbnb scores across Edinburgh, ordered by the median value of their review scores.

- Each ridge denotes a neighbourhood, illustrating the range and density if  review  scores.

- Neighbourhoods with a pronounced and concentrated peak like `Morningside` have a higher median score. Guests tend to rate their stays with a high rating.

- However neighbourhoods with a broader ridge convey a wide variety of review ratings given by the guests. 


QUESTION 2

```{r}
library(fs)
library(janitor)
library(tidyverse)
# get a list of files with "Foreign Connected PAC" in their names 
list_of_files <- dir_ls(path = "data", regexp = "Foreign Connected PAC")
```

```{r}
pac <- read_csv(list_of_files, id = "year") 
```

```{r}
pac <- pac %>%
  clean_names()

glimpse(pac)
```
```{r}
pac_separate<- pac%>%
  separate(col= country_of_origin_parent_company,
           into= c("country_of_origin","parent_company"),
       sep="/")
```
```{r}
pac_transform <- pac_separate %>%
  select(year, pac_name_affiliate, country_of_origin, parent_company, dems, repubs)%>%
  mutate(
    year = as.integer(str_extract(year, "\\d{4}(?=\\.csv)")),
    dems = as.numeric(gsub("[^0-9]", "", dems)), 
    repubs = as.numeric(gsub("[^0-9]", "", repubs))
  )
glimpse(pac_transform)
```
```{r}
pac_long <- pac_transform %>%
  pivot_longer(cols = c(dems, repubs), 
               names_to = "party", 
               values_to = "amount")
glimpse(pac_long)

```

```{r}
pac_longer <- pac_long %>%
  mutate(party = ifelse(party == "dems", "Democrat", "Republican"))

# Glimpse at the transformed data
glimpse(pac_longer)

```

```{r}
uk_pac_contributions <- pac_longer %>%
  filter(country_of_origin == "UK") %>%  # Filter for PACs from the UK
  group_by(year, party) %>%  # Group by year and party (dems/repubs)
  summarize(total_contributions = sum(amount, na.rm = TRUE), .groups = 'drop')

# View data
glimpse(uk_pac_contributions)

```

```{r}
ggplot(uk_pac_contributions, aes(x = year, y = total_contributions, color = party, group = party)) +
  geom_line(size = 1) + 
  scale_y_continuous(labels = scales::dollar_format(scale = 1e-6, suffix="M"))+
  labs(
    title = "Contributions to US Political Parties from UK_Connected PAC",
    x = "Year",
    y = "Total Amount",
    color = "Party",
    caption = "Source: OpenSecrets.org"
  ) +
  scale_color_manual(values = c("Democrat" = "blue", "Republican" = "red")) + 
  theme_minimal()+
  theme(
    plot.title = element_text(hjust = 0, size = 12),
    plot.caption = element_text(hjust = 1, face ="italic"),
    axis.title.y = element_text(hjust = 0), 
    axis.title.x = element_text(hjust = 0),
    legend.position = c(0.85,0.16)
    )

```

```{r}
switz_pac_contributions <- pac_longer %>%
  filter(country_of_origin == "Switzerland") %>%  
  group_by(year, party) %>%  # Group by year and party (dems/repubs)
  summarize(total_contributions = sum(amount, na.rm = TRUE), .groups = 'drop')
glimpse(uk_pac_contributions)

```

```{r}
ggplot(switz_pac_contributions, aes(x = year, y = total_contributions, color = party, group = party)) +
  geom_line(size = 1) + 
  scale_y_continuous(labels = scales::dollar_format(scale = 1e-6, suffix="M"))+
  labs(
    title = "Contributions to US Political Parties from Switzerland_Connected PAC",
    x = "Year",
    y = "Total Amount",
    color = "Party",
    caption = "Source: OpenSecrets.org"
  ) +
  scale_color_manual(values = c("Democrat" = "blue", "Republican" = "red")) + 
  theme_minimal()+
  theme(
    plot.title = element_text(hjust = 0, size = 12),
    plot.caption = element_text(hjust = 1, face ="italic"),
    axis.title.y = element_text(hjust = 0), 
    axis.title.x = element_text(hjust = 0),
    legend.position = c(0.85,0.16)
    )
```



QUESTION 3

```{r}
med_price <- read_csv("C:/Users/ual-laptop/Downloads/INFO526 HW2/data/median-housing.csv")
recession <- read_csv("C:/Users/ual-laptop/Downloads/INFO526 HW2/data/recessions.csv")
```


-   Rename Variables

```{r}
med_price %>%
    rename(
      date = DATE,
      price = MSPUS
    )
```

- Conversion

```{r}
mp <- med_price %>%
  mutate(new_date = as.Date(DATE, format = "%Y-%m-%d"))
```


-   Visualization 1

```{r}
ggplot(mp, aes(x= new_date, y = MSPUS))+
  geom_line(color = "blue")+
  scale_x_date(date_breaks = "5 years", date_labels = "%Y")+
  scale_y_continuous(
    breaks = seq(0, 400000, by = 40000),
    labels = scales::comma)+
  labs(title = "Median sales prices of houses sold in the United States",
       subtitle = "Not seasonally adjusted",
       caption = "Sources: Census; HUD",
       y = "Dollars",
       x = "")+
  theme_minimal()+
  theme(
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.y = element_line()
  )
  
```
- Identifying Recession

```{r}
frame_start <- min(med_price$DATE)
frame_end <- max(med_price$DATE)

id_rec <- recession %>%
  mutate(id = (Peak >= frame_start & Trough <= frame_end)) %>%
  filter(id)

id_rec$Peak <- as.Date(id_rec$Peak)
id_rec$Trough <- as.Date(id_rec$Trough)
```

- Visualization 2

```{r}
ggplot()+
  geom_rect(data = id_rec, aes(xmin= Trough, xmax = Peak, ymin=0, ymax = Inf),
              fill = "gray", alpha = 0.4)+
  geom_line(data =mp, aes(x= new_date, y = MSPUS), color = "blue")+
  scale_x_date(date_breaks = "5 years", date_labels = "%Y")+
  scale_y_continuous(
    breaks = seq(0, 400000, by = 40000),
    labels = scales::comma)+
  labs(title = "Median sales prices of houses sold in the United States",
       subtitle = "Not seasonally adjusted",
       caption = "Shaded areas indicate US recessions\n
       Sources: Census; HUD",
       y = "Dollars",
       x = "")+
  theme_minimal()+
  theme(
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.y = element_line()
  )

```

- Creating a subset

```{r}
med_sub <- mp %>%
  filter((year(new_date) == 2019) | (year(new_date) == 2020))
```

- Adding columns

```{r}
med_sub <- med_sub %>%
  mutate(year = year(new_date),
         quarter = paste0("Q",quarter(new_date)))
```

- Visualization 3

```{r}
med_sub |>
  mutate(year_quarters = factor(paste(year, "|", quarter))) |>
  ggplot(aes(x = new_date, y = MSPUS)) +
  geom_line(colour = "blue", size = 0.8) +
  geom_point(colour = "blue", fill = "white", shape = 21) +
  scale_y_continuous(name = "Dollars",
                     breaks = seq(300000, 360000, by = 20000),
                     labels = scales::comma,
                     ) +
  scale_x_date(name = "",  
               breaks = med_sub$new_date,
               labels = med_sub$quarter) +  
  labs(title = "Median sales price of houses sold in the United States",
       subtitle = "Not seasonally adjusted", x=NULL) +
  coord_cartesian(clip = "off", ylim = c(300000, 360000),
                  xlim = c(as.Date("2019-01-01"), as.Date("2020-12-31"))) +
  annotate("text", x = as.Date("2019-06-01"), y = 290000, label = "2019") +
  annotate("text", x = as.Date("2020-06-01"), y = 290000, label = "2020") +
  theme_minimal() +
  theme(plot.title.position="plot")
```

- QUESTION 4

```{r}
# Load necessary libraries
library(ggplot2)
library(ggforce)

# Create a data frame for circles
target_data <- tibble(
  x = 0, 
  y = 0, 
  r = c(1, 0.70, 0.40)  # Radii for each circle
)

# Create the target symbol
ggplot(target_data) +
  geom_circle(aes(x0 = x, y0 = y, r = r), 
              color = "#cc0000", 
              fill = "#cc0000") +  # Outer circle
  geom_circle(aes(x0 = x, y0 = y, r = r[2]), 
              color = "white", 
              fill = "white") +  # Inner circle (white)
  geom_circle(aes(x0 = x, y0 = y, r = r[3]), 
              color = "#cc0000", 
              fill = "#cc0000") +  # Middle circle (red)
  coord_fixed() +  # Keep aspect ratio
  theme_void() +   # Remove axes and background
  annotate("text",x= 0, y= -1.2,label = "TARGET", color = "#cc0000", size= 8, fontface="bold")+
  annotate("text", x=0.4, y= -1.25, label="®", color ="#cc0000", size= 5, fontface="bold")
```


QUESTION 5

```{r}
library("palmerpenguins")

ggplot(penguins, aes(x= bill_length_mm, y= bill_depth_mm))+
  geom_point(aes(color = species, shape = species), size = 2.5)+
  labs(x= "Bill length",
       y= "Bill depth",
       title = "Bill depth VS bill length of various penguin species")
```


```{r}
ggplot(penguins, aes(x= bill_length_mm, y= bill_depth_mm, color= species, shape = species))+
  geom_point(size = 2.5)+
  scale_color_manual(values = c("Adelie"="#15F5BA", "Chinstrap"="#97FFF4", "Gentoo" = "#C0E218"))+
  theme(
    panel.background = element_rect(fill = "#FCF876"),
    plot.background = element_rect(fill = "#851DE0"),
    legend.background = element_rect(fill = "#851DE0")
  )
```

- For the sake of this exercise I have created this plot with a very data-to-ink ratio, which is considered to be ugly.


```{r}
ggplot(penguins)+ xlim(0, 10) + ylim(0, 100)
```

- The reason for this empty plot is that, for me it is truly ugly as there is no data in this. Even in graphs with high data-to-ink ratios you can get something, but here from a data scientist's perspective this would be the ugliest plot.
